\subsection{Fréchet spaces and algebras}
%---------------------------------------

\begin{definition}
A \defe{Fréchet space}{Fréchet!space} is an Hausdorff locally convex space whose topology can be defined by a \emph{countable} family of seminorm and which is complete for any distance invariant under the translations.
\end{definition}

Somme results without proof :
\begin{lemma}\dref{12.4.6}
 An Hausdorff uniformisable space whose topology can be defined by a sequence $(d_n)$ of \ecarts\ is metrisable.
\end{lemma}

\begin{lemma}\dref{12.14.5}
An Hausdorff locally convex space whose topology can be defined by a \emph{countable} family of seminorms is metrisable.
\end{lemma}

In particular, a Fréchet space is metrisable.

\begin{lemma} \label{prop_suiteFk}
Let $(F_k)$ be a decreasing sequence of Fréchet spaces such that the canonical injections $\dpt{\id}{F_{K+1}}{F_k}$ are continuous. Let, for each $k\in\eN$, be a sequence $(a_{k,q})$ of points in $F_k$ which converges to zero in the sense of $F_k$. Then there exists an increasing sequence $q_k\to\infty$ such that for all $h\geq 0$, 
\[ 
  \sum_{k\geq h}a_{k,q_k}
\]
is a converging sum in $F_k$.
\end{lemma}

The following result is easy to prove :

\begin{proposition}
If $A$ is a Fréchet algebra for the seminorms $\| . \|$, the space $C_{\flat}(V,A)$ is a Fréchet algebra for the seminorms 
\[ 
  \| f \|_k=\sum_{v\in V}\| f(v) \|_k
\]
\end{proposition}


\begin{definition}
A \defe{Fréchet algebra}{Fréchet!algebra} is a topological algebra whose topology is given by a countable family of sub-multiplicative seminorms :
\[ 
  \| ab \|_k\leq \| a \|_k\| b \|_k.
\]

\end{definition}

\begin{proposition}
Let $A$ be a Fréchet algebra with seminorms $\| . \|_k$ and $\alpha$ be an action of the vector space $V$ on $A$. We put on $A^{\infty}$ the seminorms
\begin{equation}
  \| a \|_{jk}=\frac{1}{ \mu! }\sup_{i\leq j}\sum_{| \mu |\leq k}\| \partial_{\mu}a \|_i.
\end{equation}
These seminorms are sub-multiplicative.

\end{proposition}

\begin{proof}
The element $\partial_{\mu}a\in A^{\infty}$ is defined by the action: for each $X\in V$ we set
\[ 
  \partial_X a=\Dsdd{ \alpha_{tX}(a) }{t}{0}
\]
and $\mu$ is a multi-index. The sum $\sum_{| \mu |\leq k}\| \partial_{\mu}(ab) \|_{i}$ is made of terms of the form
\[ 
  \| \partial^{\alpha}a\partial^{\beta}b \|_i\leq\| \partial^{\alpha}a \|_i\| \partial^{\beta} \|_i.
\]
Evidently, $\sup_{i\leq j}\| \partial^{\alpha}a \|_i\leq \| a \|_{jk}$. So $\| ab \|_{jk}$ is made of $\mu !$ terms which are all lesser than $\| a \|_{jk}\| b \|_{jk}$. The factor $1/\mu!$ concludes the proof.

\end{proof}


